### 2010-06-26 paper/IIWAS2010/content/system_arch.tex
Changes after Friday's meeting

For ontology construction Morpheus uses the Wikipedia pages, categories,
and the WordNet synset heterarchy. This approch is motivated from
YAGO\cite{Suchanek2009phd}. For a given realm, we first find a mapping category 
from the DBpedia categories\cite{Auer07dbpedia:a} (these are extracted from 
the Wikipedia pages). Then we grab all the neighboring categories
using a \textit{Markov Blanket} \cite{PRIS}. In fact, we grab 
the category's parents, its children, and its 
children's other parents using the DBpedia ontology properties \textit{broader}
and \textit{narrower}. Once the blanket is ready, it grabs all parent categories
of the categories in the blanket till it reaches root. Once we have the
categories belongs to a particular domain, we associate them with WordNet synsets
using an algorithm similar to YAGO\cite{Suchanek2009phd}. We determine head noun in the
category name and search for it in the WordNet synsets.
Those Wikipedia categories having no WordNet match are ignored in the our realm
ontology. In fact we uses the WordNet synset heterarchy as the backbone of the   
realm ontology.

To build a corpus for each of the leaf nodes in the ontology, 
we extract \emph{n-grams} or \emph{terms} from the Wikipedia pages associated with the 
the DBpedia categories \cite{Auer07dbpedia:a}. From this category-term corpora, 
we can find out the likelihood that a term belongs to a particular category, 
which is useful in categorizing the terms in a \emph{regular user's} query [section \ref{sec:qrm_ranking}].    
This likelihood is determined by the probability of a category given a term 
using \textit{Bayes Rule} (Eq. \ref{eq:bayesrule}), since we can 
easily obtain the term-category and term-corpus probabilities as 
relative frequencies. The higher the probability, the more likely a 
term is referring to a category.

@@

\subsection{Topic Models and Dcoument Classification }
\label{sec:topicmodels}

@@

To answer a regular user's query, a \textit{candidate
SSQ}, Morpheus finds similar SSQs that belong to QRMs the Morpheus data store (a
\textit{qualified SSQ}). We need a similarity measure to match the candidate SSQ
with a qualified SSQ. For the SSQ similarity, we
consider the SSQ components such as query-realm, input
terms and output terms, and their assigned categories or classes. The,
\textit{category divergence} of two categories characteristics their 
similarity based on an ontology class heterarchy. 
It is motivated by the concept of multiple
dispatch in CLOS \cite{Steele1990} and Dylan programming \cite{Barrett1996} for
generic function type matches.

We use \emph{category divergence} to calculate the relevance between the candiate SSQ
and a qualified SSQ. Each qualified SSQ will
have input terms, output terms, and associated classes, and one realm. For the
candidate SSQ, the relevant category for input terms are determined from the
ontology corpora [section \ref{sec:ontology_corpora}]. The relevance 
of a qualified SSQ to the candidate SSQ is determined by
aggregating the divergence measure of categories associated with them. In addition, 
We order QRMs in the data store in order by descreasing relevance. 
The order provides a ranking for the results to the user. 
The following subsection \ref{sec:ctd} describe \emph{category divergence} in detail.

@@

We employ a quasimetric, \textit{category
divergence},
between a source category and a target category using the \textit{topological
structure} of the categories in an ontology. We write $S \prec T$ for the
reflexive
transitive closure of the supercategory relation. Let $d(P,Q)$ represent the hop
distance in the directed ontology inheritance graph from $P$ to $Q$. The
divergence between a source and target category ranges from zero (for identical
categories) to one (for incompatible categories). Let $S$ be the source
category, $T$ be
the target category, and let $C$ be a least common ancestor category of $S$ and
$T$
minimizing $d(S,C) + d(T,C)$. The category divergence between $S$ and $T$ is
defined given by:

\begin{equation}
d_{cat}(S, T) = \begin{cases}
0 & S.{Uri} \equiv T.{Uri}\\
d(S, T)/(3h) & S \prec T\\
1 & T \prec S\\
(d(S,root) + d(S,C) \\ \ \ \ \ + d(T,C))/(3h) & otherwise
\end{cases}
\end{equation}

@@

Note, if $S \prec T$ and $S \not\prec Q$ then $d_{cat}(S,T) <
d_{cat}(S,Q)$, that is, the divergence of a source category to a target
ancestor category is smaller than the divergence of a source category to any
category that is not an ancestor. This is an important property in
determining the compatibility of category for answering queries.  If a
SSQ answers queries concerning an ancestor category, it is more relevant
that a SSQ that answers queries from any non-ancestral category.

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=90mm]{img/automotive_ontology.eps}
\caption{An abbreviated graphical representation of an automotive
ontology: This is a manually created ontology, from the Wikipedia category
hierarchy and pages. The edges represent inheritance relationships between the
category in a topology.}
\label{fig:automotive_ontology}
\end{figure}

Suppose we want to find the category divergence between \textit{Toyota Camry}
and \textit{Coupes} [Figure \ref{fig:automotive_ontology}]. 
\textit{Toyota Camry} is the subcategory of \textit{Sedans}, and \textit{Coupes}
is the subcategory of \textit{Cars}. Therefore, the least common ancestor ($C$)
for these two categories is \textit{Cars}. In addition, the tree height is $5$.
So, the normalized divergence $d_{cat}$\textit{(Toyota Camry, Coupes)} is $7/15$
that is calculated from \textit{d(Toyota Camry, Root)} is 5, \textit{d(Toyota
Camry, Cars)} is two, and \textit{d(Coupes, Cars)} is one.

>>>>

For ontology construction Morpheus uses the Wikipedia pages, categories,
and the WordNet synset heterarchy. This approch is motivated from
YAGO\cite{Suchanek2009phd}. For a given realm, we first find a mapping category 
from the DBpedia categories, which are extracted from 
the Wikipedia pages\cite{Bizer2009}. Then we grab all the neighboring categories
using a \textit{Markov Blanket} \cite{PRIS}. In fact, we grab 
the category's parents, its children, and its 
children's other parents using the DBpedia ontology properties \textit{broader}
and \textit{narrower}. Once the blanket is ready, it grabs all parent categories
of the categories in the blanket till it reaches root. Once we have the
categories belongs to a particular domain, we associate them with WordNet synsets
using an algorithm similar to YAGO\cite{Suchanek2009phd}. We determine head noun in the
category name and search for it in the WordNet synsets.
Those Wikipedia categories having no WordNet match are ignored in the our realm
ontology. In fact we uses the WordNet synset heterarchy as the backbone of the   
realm ontology.

To build a corpus for each of the leaf nodes in the ontology, 
we extract \emph{n-grams} or \emph{terms} from the Wikipedia pages associated with the 
the DBpedia categories\cite{Bizer2009}. From this category-term corpora, 
we can find out the likelihood that a term belongs to a particular category, 
which is useful in categorizing the terms in a \emph{regular user's} query [section \ref{sec:qrm_ranking}].    
This likelihood is determined by the probability of a category given a term 
using \textit{Bayes Rule} (Eq. \ref{eq:bayesrule}), since we can 
easily obtain the term-category and term-corpus probabilities as 
relative frequencies. The higher the probability, the more likely a 
term is referring to a category.

@@

@@

To answer a regular user's query, a \textit{candidate
SSQ}, Morpheus finds similar SSQs that belong to QRMs the Morpheus data store (a
\textit{qualified SSQ}). We need a similarity measure to match the candidate SSQ
with a qualified SSQ. For the SSQ similarity, we
consider the SSQ components such as query-realm, input
terms and output terms, and their assigned categories or classes. The,
\textit{category divergence} of two categories characteristics their 
similarity based on an ontology class heterarchy. 
It is motivated by the concept of multiple
dispatch in CLOS and Dylan programming for generic function \emph{type} matches.

We consider the category match as a type match and we use 
\emph{category divergence} to calculate the relevance between 
the candiate SSQ and a qualified SSQ. Each qualified SSQ will 
have input terms, output terms, and associated classes, and one realm. 
For the candidate SSQ, the relevant category for input terms 
are determined from the ontology corpora [section \ref{sec:ontology_corpora}]. The relevance of a qualified SSQ to the candidate SSQ is 
determined by aggregating the divergence measure of categories 
associated with them. In addition, we order QRMs in the data store 
by descreasing relevance. The order provides a ranking for the results 
to the user. The following subsection describes \emph{category divergence} 
in detail.

@@

We employ a quasimetric, \textit{category
divergence (cd)},
between a source category and a target category using the \textit{topological
structure} of the categories in an ontology. We write $S \prec T$ for the
reflexive
transitive closure of the supercategory relation. Let $d(P,Q)$ represent the hop
distance in the directed ontology inheritance graph from $P$ to $Q$. The
divergence between a source and target category ranges from zero (for identical
categories) to one (for type incompatible categories). Let $S$ be the source
category, $T$ be
the target category, and $C$ be a least common ancestor category of $S$ and
$T$ i.e., one that minimizes $d(S,C) + d(T,C)$. The category divergence between $S$ and $T$ is defined given by:

\begin{equation}
cd(S, T) = \begin{cases}
0 & S.{Uri} \equiv T.{Uri}\\
d(S, T)/(3h) & S \prec T\\
1 & T \prec S\\
(d(S,root) + d(S,C) \\ \ \ \ \ + d(T,C))/(3h) & otherwise
\end{cases}
\end{equation}

@@

Note, if $S \prec T$ and $S \not\prec Q$ then $cd(S,T) <
cd(S,Q)$, that is, the divergence of a source category to a target
ancestor category is smaller than the divergence of a source category to any
category that is not an ancestor. This is an important property in
determining the compatibility of categories for answering queries.  If a
SSQ answers queries concerning an ancestor category, it is more relevant
that a SSQ that answers queries from any non-ancestral category.

% Algorithm 1: figure  
\begin{figure}[t]
\centering
\includegraphics[width=85mm]{img/automotive_ontology.eps}
\caption{An abbreviated man-made automotive ontology: Arrows represent inheritance relationships between the ontological classes.}
\label{fig:automotive_ontology}
\end{figure}

Suppose we want to find the category divergence between \textit{Sedans}
and \textit{Trucks} [Figure \ref{fig:automotive_ontology}]. 
\textit{Sedans} is the subcategory of \textit{Cars}, and \textit{Trucks}
is the subcategory of \textit{Automobiles}. Therefore, the least common ancestor ($C$)
for these two categories is \textit{Automobiles}. In addition, the tree height \textit{h = 4}.
So, the normalized divergence $cd$\textit{(Sedans, Trucks)} is \textit{7/12}
that is calculated from \textit{d(Sedans, Root) = 4}, \textit{d(Sedans, Automobiles) = 2}, and \textit{d(Trucks, Automobiles) = 1}.
